\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-08-26 ResponseToReview.tex

\mbh{I agree with Jeff here}

%JRI STOPPED HERE

11) Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.\\

\res{The sizes of amplicons are now mentioned explicitly in the Materials and Methods (page \mbh{X}, lines \mbh{XX}) following primer sequences: 
``Homozygotes show a single band for absence of the \emph{Hopscotch} element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."
The 5kb amplicon including the entire \emph{Hopscotch} could be amplified in a single reaction.}

@@

12) Page 6 line 6: "When only one PCR resolved well, we scored one allele for the individual".  Sorry, what does this mean? That you have one band (one assumes the short one?). If you are talking presence/absence, how do you know the allele?\\

\res{We have clarified this in the text on page \mbh{X}, lines \mbh{XX}.  We developed primers for two PCR reactions that would amplify products with size polymorphism characteristic of the two different alleles (\emph{i.e.}, \emph{Hopscotch} vs. no \emph{Hopscotch}). If PCR products were clearly resolved for only one of these reactions, we included one allele for that individual in our data set rather than infer the diploid genotype. We have included the following sentence in the text: \mbh{update the following in the text} ``Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}

@@

13) Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include?\\

\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page \mbh{X}, line \mbh{XX}: \mbh{Update text in manuscript}``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}

@@

14) Page 7 lines 17 - 26: how many year's worth of environmental data went into your analysis? Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?\\

%\res{The environmental data used from Pyh\"aj\"arvi  \emph{et al.} (2013) was one year's worth of environmental data. A previous study by Hufford \emph{et al.} (2012) has shown that environmental conditions in areas where teosinte grows have not changed significantly over the last 10,000 years. We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database (FAOHWSD) and \url{www.harvestchoice.org} and summarized by principle component analysis following Pyh\"aj\"arvi  \emph{et al.} (2013)."}

\res{Our climatic data consisted of the 19 bioclim layers commonly used in ecological niche analyses.  Full methods for generation of these data can be found at the following url: \url{http://www.worldclim.org/methods}.  Briefly, temperature and precipitation were averaged for each global location over a minimum of ten years during the period from 1950-2000.  Soil data were downloaded from the Harmonized World Soil Database (FAOHWSD; \url{http://webarchive.iiasa.ac.at/Research/LUC/External-World-soil-database/HWSD_Documentation.pdf}). This recently published data set drew together global soil maps generated by a world-wide network of scientists over the last several decades and summarized them in GIS data layers that are available as polygon coverages. Data for our sampling sites were extracted from these data sets based on latitude and longitude using the DIVA-GIS platform.  Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.  We have further clarified this on page \mbh{X}, lines \mbh{XX} of the text.}

### 2014-08-26 ResponseToReview.tex

This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\

Methods and statistics were appropriately used as far as my expertise allows me to judge.
Specific suggestions follow:

Introduction:
Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".  Can you tell us why an understanding is useful? Can you give some examples?\\

\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice (Kovach and McCouch 2008). Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize (Gallavotti \emph{et al.} 2004; Sigmon and Vollbrecht 2010). While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.} Bai and Lindhout 2007), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors (Whitton \emph{et al.} 1997).}

### 2014-08-26 manuscript.tex
Edits to introduction
Response to reviewer 2, additional teosinte examples in introduction

Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}. Plant domestication usually involves a suite of phenotypic changes, such as increase yield, loss of seed shattering, and increase fruit or grain size, which are commonly referred to as 'domestication syndrome' \citep{OlsenWendel2013}.  Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}. For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}. \lev{teosinte example here} While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.

>>>>

Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}. Plant domestication usually involves a suite of phenotypic changes, such as increase yield, loss of seed shattering, and increase fruit or grain size, which are commonly referred to as 'domestication syndrome' \citep{OlsenWendel2013}.  Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}. For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}. Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrect2010}. While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
